2.4 OLS: Precision and Diagnostics - R Practice

### 2019-10-03 static/homeworks/03-problem-set-answers.Rmd
Add 12 class content

1.

a. In your own words, explain the difference between endogeneity and exogeneity.^[Hint: you may want to look back at class 1.2 for this question.]
b. In your own words, explain how conducting a randomized controlled trial helps to identify the causal connection between two variables.

\clearpage

2.

a. In your own words, explain what (sample) standard deviation *means*.
b. In your own words, explain how (sample) standard deviation *is calculated.* You may also write the formula, but it is not necessary.

\clearpage

>>>>

## Question 1

### Part A
**In your own words, explain the difference between endogeneity and exogeneity.**^[Hint: you may want to look back at class 1.2 for this question.]

The $R^2$ is a measure of how well the OLS regression line "fits" our observed data points. It is a measure of the share of the total variation in $Y$ (TSS) that is explained by the variation from our model (ESS), where:

$$R^2=\frac{ESS}{TSS} = \frac{\displaystyle \sum^n_{i=1} (\hat{Y_i}-\bar{Y})^2}{\displaystyle \sum^n_{i=1}(Y_i-\bar{Y})^2}$$

Equivalently, you can also estimate $R^2$ as:

$$R^2=1-\frac{SSE}{TSS} = 1 - \left(\frac{\displaystyle \sum^n_{i=1} \hat{u_i}^2}{\displaystyle \sum^n_{i=1}(Y_i-\bar{Y})^2}\right)$$

or, most simply:

$$R^2=[cor(X,Y)]^2$$

The closer $R^2$ is to 1, the better the fit, the closer to 0, the poorer the fit. Low $R^2$ tells us that there are better models, including more variables, that explain the variation in Y.

### Part B

**In your own words, explain how conducting a randomized controlled trial helps to identify the causal connection between two variables.**

## 2.

### a.

In your own words, explain what (sample) standard deviation *means*.

### b.

In your own words, explain how (sample) standard deviation *is calculated.* You may also write the formula, but it is not necessary.

### 2019-10-03 static/r/11-r-practice-rmd.Rmd
Add 12 class content

---
title: "2.4 OLS: Precision and Diagnostics - R Practice (Answers)"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

Our [data](https://github.com/fivethirtyeight/data/tree/master/congress-trump-score) comes from fivethirtyeight's [Trump Congress tracker](https://projects.fivethirtyeight.com/congress-trump-score/).

- [View Project on R Studio Cloud](https://rstudio.cloud/project/575252)

## Question 1
Download and read in (`read_csv`) the data.

- [<i class="fas fa-table"></i> `econfreedom.csv`](/data/congress-trump-score.csv)

@@

@@

### Part C
Look in your original data to match this outlier with an observation. Hint: use the `slice()` command, as the outlier test gave you an observation number!

@@

This data is still a bit messy. We'd also probably like to plot our scatterplots with colors for Republican and Democratic party. Or sort by the House and the Senate.

>>>>

---
title: "2.4 OLS: Precision and Diagnostics - R Practice"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    #theme: 
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up
To minimize confusion, I suggest creating a new `R Project` (e.g. `regression_practice`) and storing any data in that folder on your computer.

Alternatively, I have made a project in R Studio Cloud that you can use (and not worry about trading room computer limitations), with the data already inside (you will still need to assign it to an object).

- [<i class="fas fa-cloud"></i> View Project on R Studio Cloud](https://rstudio.cloud/project/575252)

## Question 1
Our [data](https://github.com/fivethirtyeight/data/tree/master/congress-trump-score) comes from fivethirtyeight's [Trump Congress tracker](https://projects.fivethirtyeight.com/congress-trump-score/). Download and read in (`read_csv`) the data.

- [<i class="fas fa-table"></i> `congress-trump-score.csv`](/data/congress-trump-score.csv)

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

@@

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

@@

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C
Look in your original data to match this outlier with an observation. Hint: use the `slice()` command, as the outlier test gave you an observation (row) number!

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

@@

This data is still a bit messy. Let's check in on your `tidyverse` skills again! For example, we'd probably like to plot our scatterplots with colors for Republican and Democratic party. Or plot by the House and the Senate.

### Part A
First, the variable `congress` (session of Congress) seems a bit off. Get a `count()` of `congress`.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B
Let's get rid of the `0` values for `congress` (someone made a mistake coding this, probably). Also, while we're at it, let's take `agree_pct` and `mutate` a variable that is a proper percentage (i.e. `*100`).

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C
The variable `party` is also quite a mess. `count()` by `party` to see. Then let's `mutate` a variable to make a better measure of political party - just `"Republican"`, `"Democrat"`, and `"Independent"`. Try doing this with the `case_when()` command (as your `mutate` formula).^[The syntax for `case_when()` is to have a series of `condition ~ "Outcome"`, separated by commas. For example, one condition is to assign both `"Democrat"` and `"D"` to `"Democrat"`, as in `party %in% c("Democrat", "D") ~ "Democrat"`. You could also do this with a few `ifelse()` commands, but that's a bit more awkward.] When you're done `count()` by your new party variable to make sure it worked.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part D
Now plot a scatterplot (same as Question 3) and set `color` to your party variable. Notice `R` uses its own default colors, which don't match to the actual colors these political parties use! Make a vector where you define the party colors as follows: `party_colors <- c("Democrat" = "blue", "Republican" = "red", "Independent" = "gray")`. Then, run your plot again, adding the following line to customize the colors `+scale_colour_manual("Parties", values = party_colors)`.^[`"Parties"` is the title that will show up on the legend, feel free to edit it, or remove the legend with another layer `+guides(color = F)`.]

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part E
Now facet your scatterplot by `chamber`.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---
